### 2017-08-23 figures.Rmd
Update figures Rmd; captions written; more references

---
title: "On the minimum leaf conductance"
author: ""
date: "1 August 2017"
output: 
  html_document:
    theme: flatly
csl: global-change-biology.csl
bibliography: references.bib
---

```{r, echo=FALSE}
source("R/load.R")
pacman::p_load(magicaxis)
knitr::opts_chunk$set(warning = FALSE, echo=FALSE)
```

```{r, fig.width=6, fig.height=6}
figure_sim()
```

**Figure 1**. Simulations  of intrinsic water-use efficiency (A~n~/g~s~), intercellular CO~2~ concentration (C~i~) and leaf transpiration rate (E~L~) with three values of g~0~. For the top two panels, all environmental drivers were held constant except the photosynthetic photon flux density (PPFD), demonstrating that the inclusion of g~0~ is necessary to simulate the decrease in A~n~/g~s~ at low PPFD. For the bottom two panels, VPD and T~air~ were covaried based on an empirical relationship [@duursma2014], reflecting typical covariation in field conditions. For all simulations it is assumed that T~leaf~ is equal to T~air~.

```{r, fig.width=8, fig.height=4}
figure_g0g1_cor(lin2015, "Nicolas Martin-StPaul_Quercus ilex_StPaul_Puechabon")
```

**Figure 2**. Example linear regression of stomatal conductance (g~s~) against a stomatal index, in this case the linearized version of the model of Medlyn et al. (2011). This example shows a very good fit between g~s~ and the stomatal index, and was selected from the database of Lin et al. (2015). The solid line is regression line, and the shaded area a 95% confidence interval for the mean. The right panel visualizes the correlation between estimated slope (g~1~) and intercept (g~0~) of the regression. The dotted ellipse is a bivariate 95% confidence interval for slope and intercept. The symbols represent 1000 bootstrap samples of the coefficients.

```{r, fig.width=9, fig.height=4}
figure_R2g0(lin2015coef, miner)
```

**Figure 3**. Estimates of g~0~ depend on the overall goodness of fit. In the left panel, we fitted the linearized form of the Medlyn et al. (2011) model to each of the datasets in the Lin et al. (2015) database. Vertical lines are 95% confidence intervals. In the middle panel, a very similar association between g~0~ and R^2^ is presented by the review of @miner2017, using the Ball-Berry model. In the right panel, shown is the correlation between the standard error of the estimate of g~0~ and the coefficient of variation of the Ball-Berry coefficient (A/Ca*sqrt(D)), showing that good estimates of g~0~ can only be obtained when there is much variation in the drivers.

```{r, fig.width=5, fig.height=5}
figure_gmin_review_2(gdfr)
```

**Figure 4**. Comparison of various estimates of the minimum conductance. Error bars are 95\% confidence intervals. Different letters denote significant differences (at $\alpha$ = 0.05). See Table 1 for definitions of minimum conductance, and text for the data sources. g~cut,isol~, cuticular permeance on isolated cuticles, g~cut,seal~, mass loss of detached leaves when stomatal side of the leaf is sealed (*might remove this*), g~min~ minimum conductance with mass loss of detached leaves, g~dark~ nighttime conductance (from Lombardozzi), g~0~ fitted parameter, using the Lin database and Miner 2017, g~s,lowA~ stomatal conductance from the Lin database, calculated for each study where A~n~ < 2.

```{r, fig.height=3, fig.with=8}
gmin_3panel(gmindat, cropgmin)
```

**Figure 5**. Literature compilation of minimum conductance (g~min~), measured with mass loss of detached leaves. Left panel, distribution across all 220 species values extracted from the literature, showing a log-normal distribution. Mean=, sd=. Middle panel, g~min~ shown by taxonomic order. Note higher g~min~ for grasses (Poales), but otherwise few clear differences. Sample size (number of species) shown on top of the figure. Right panel: genotypic variation in g~min~ for 9 major crops. Sample sizes are the number of distinct crop genotypes on which g~min~ was reported for each crop, bars denote the range in g~min~ across all the data for a crop, illustrating the very large genotypic variation compared to inter-specific variation.

```{r, fig.width=3.5, fig.height=4 }
figure_wtc4_gmin_2(wtc4gmin)
```

**Figure 6**. Minimum conductance measured from leaf dry-down curves on *Eucalyptus parramattensis* grown in whole-tree chambers. Trees were grown following ambient conditions, or in a constant elevated temperature treatment (+3 degrees C). Measurements of g~min~ were carried out at five different temperatures on replicate leaves. There is no effect of measurement T on g~min~, but leaves of trees grown in elevated temperature showed consistently lower g~min~.

```{r, fig.width=6, fig.height=5}
figure_hakea_gmin(lopw)
```

**Figure 7**. Minimum conductance measured from leaf dry-down curves on a variety *Hakea* species, a genus native to Australia. Plants were grown in containers in a grow house, and supplied with ample water or underwent long-term mild drought stress. Bars are labelled by leaf form : broadleaf (B) or needle-like (N). For droughted plants, g~min~ was higher for species with needle-like leaves (Wilcoxon test, p `r pval(wilcox.test(gmin.d ~ leaf.form, data=lopw)$p.value)`)), but not for well-watered plants (p `r pval(wilcox.test(gmin.w ~ leaf.form, data=lopw)$p.value)`).

# References