### 2018-01-11 manuscript_suppinfo.Rmd
moved table 2 to suppinfo; suppinfo with references and two plots

---
title: "Supporting Information"
author: ''
date: ''
output:
  word_document:
    reference_docx: manuscriptstyle.docx
csl: new-phytologist.csl
bibliography: references.bib
---

```{r include=FALSE}

library(knitr)
library(pander)

source("R/load.R")
```

# 1. Literature compilation of g~min~ estimates

\
Data on minimum conductance (g~min~) from MLD was compiled from 40 original papers. For the compilation, we returned to all original papers mentioned in the review by @kerstiens1996, and added many newer sources. All conductance values were converted to per unit projected surface area. Total surface area for conifer leaves was converted to projected area following @barclay2000. Units were converted to molar fluxes assuming 41 mol m^-3^ (cf. @kerstiens1996). In many cases, minimum transpiration rates were reported instead of a conductance (termed 'residual transpiration', 'epidermal transpiration', 'minimum transpiration', etc.). For all papers included in the compilation, we were able to convert these transpiration rates to conductances, but unfortunately dozens of papers could not be included because of missing information on vapour pressure deficit during measurements. All g~min~ estimates from original papers were entered in the database, but for this paper we only used 'control' treatments (filtering for drought, elevated temperature, and various other treatments). Where practical, other reported variables were also entered for each study. The complete database is available from github.com/remkoduursma/gmindatabase.

\
Taxonomic order and family was looked up with the taxize R package [@taxize2013]. Climate variables across the species range were extracted from WorldClim, using species occurrence records from ALA and GBIF. The speciesmap R package was used for this extraction (github.com/remkoduursma/speciesmap).

```{r gminfig1, echo=FALSE, dpi=300, fig.width=6, fig.height=5}
figure_gmin_bygroup(gmindat)
```
**Figure S1**.

```{r gminfig2, echo=FALSE, dpi=300, fig.width=6, fig.height=5}
# climate plot
```
**Figure S2**.

## References

```{r echo=FALSE, eval=FALSE}
 library(RefManageR)
r <- ReadBib("c:/repos/gmindatabase/references/references.bib")
paste(paste0("@", names(r)), collapse="; ")
```

[@anfodillo2002; @araus1991; @arnold2007; @bengtson1978; @boyce2003; @boyce2000; @brodribb2014; @burghardt2003; @cavender-bares2007; @cornwell2007; @drake2013; @vangardingen1991; @grace1974; @hamerlynck1996; @hao2010; @heinsoo1998; @holbrook1992; @iogna2013; @james2008; @jordan2007; @kerstiens1995; @kerstiens1989; @li2012; @muchow1989; @nagarajah1979; @nazrul-islam1983; @pasquet-kok2010; @phillips2010; @pisekberger1938; @rawson1988; @sack2003; @saito2010; @schuster2016; @scoffoni2011; @sinclair1986; @smith2006; @vilagrosa2003; @villar-salvador1999; @walden-coleman2013; @warren2011; @pallardy1993; @newton1996; @fish2009; @rosas-anderson2014]

# 2. Effects of drought stress on g~min~

```{r, echo=FALSE}
dat <- read.csv("data/drought_gmin_table.csv", stringsAsFactors = FALSE)

knitr::kable(dat, 
             caption="**Table S1.** Summary of studies demonstrating a decrease in g~min~ after an acclimation period to low soil water availability",
             col.names= c("Study","Species","Experiment","Reduction in g~min~"))
```

# 2. Effects of leaf age on g~min~

```{r leafagefig, echo=FALSE, dpi=300, fig.width=8, fig.height=6}
figure_leafage(gminall)
```

**Figure S3**. Correlation between g~min~ and leaf age for six studies in the g~min~ database. Leaf age is expressed in years (0 = current year), with the exception of @phillips2010, who measured juvenile and adult foliage on *Eucalyptus*. Citations with species studied are @boyce2000 (*Abies lasiocarpa*, *Picea engelmanii*), @boyce2003 (*Tsuga canadensis*, *Pinus strobus*, *Pinus resinosa*, *Picea rubens*), @heinsoo1998 (*Picea abies*), @jordan2007 (*Agastachys odorata*), @phillips2010 (*Eucalyptus saligna*, *E. sideroxylon*, *E. cladocalyx*).

# References {.unnumbered}